Simple Wind Models

### 2010-07-12 Winds/papers/Theory/First/wind_simple.tex
profx
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5035 ac11df5e-a788-c23b-a8e1-c758c323924d

\begin{itemize}
\item Usage of cool gas outflows to probe galactic-scale winds in
  star-forming galaxies is now a well-established observational
  technique
\item Simple anlaysis focusing on the central wavelength and EW of the
  lines
\item \lya, MgII, NaI and (now) FeII* show emission
\item Despite the wealth of observational diagnostics, little effor
  hast been taken to constrain the wind properties
\item These are some first steps by considering idealized wind
  profiles.  Algorithms can be extended to any wind output
\end{itemize}

>>>>

Nearly all gaseous objects in astrophysics that emit light are also
observed to drive gas away.  [word] This includes the jets of protostars, the
strong stellar winds of massive O and B stars, the gentle Solar wind
of our Sun, the associated absorption of bright quasars, and the
spectacular jets of radio-loud AGN.   These gaseous outflows moderate
the accretion of fresh material onto the object, inject energy and
momentum into gas on large scales, and ...
[Last line]

Star-burst galaxies, whose light is dominated by \ion{H}{2} regions
and massive stars, are also observed to drive gaseous outflows.  The
flows are generally expected (and sometimes observed) to have multiple
phases, with a hot and diffuse phase traced by X-ray emission
\citep[e.g.][]{strickland,martin+k} and a cool, denser phase traced by
H$\alpha$ emission \citep[e.g.][]{NGC891,M87}.

Beyond the local universe, galactic outflows are predominatly
revealted by UV and optical absorption lines, e.g.\ \ion{Na}{1},
\ion{Mg}{2}, \ion{Si}{2} and \ion{C}{4} transitions.  With the galaxy
as a backlight, one almost universally observes predominantly
blue-shifted material indicating a galactic-scale flow of gas toward
Earth and away from the galaxy.  These transitions are sensitive to
the cool (\ion{Mg}{2}; $T \sim 10^4$K) and warm (\ion{C}{4}; $T \sim
10^5$K) phases of the flow.  The incidence of cool gas outflows is
nearly universal in star-forming galaxies;  this includes systems at $z \sim 0$
which exhibit \ion{Na}{1} and \ion{Mg}{2} absorption
\citep{rupke,martin,student}, $z \sim 1$ star-forming galaxies traced by
\ion{Fe}{2} and \ion{Mg}{2} transitions \citep{wcp+09,rubin+10}, and
$z>2$ Lyman break galaxies (LBGs) that show blue-shifted \ion{Si}{2},
\ion{C}{2}, and \ion{O}{1} transitions \citep{lowenthal,steidel94}.

Although the obsevation of metal-line absorption is a well established
means of identifying outflows at all epochs, little effort has been
made to compare these observations against (even idealized) wind
models.  The focus to date has been to gelan the limited information
affored by direct analysis of the absorption lines.  The data do yield
robust measurements of the speed of the gas yet poorly constrain the
optical depth, covering fraction, density, temperature, and distance
of the gas from the emitting regions.  Constraints related to the
mass, energetics, and momentum of the flow suffer from
orders of magnitude uncertainty.  Both the origin and impact of the
winds, therefore, remain open points of debate.

Recent studes of $z \sim 1$ star-forming galaixes have revealed that
the cool gas frequently exhibits significant \ion{Mg}{2} emission in
addition to the ubiquitous blue-shifted absorption
\citep{wcp+09,rubin+10b}.  \cite{rubin+10a} have also detected line
emission from non-resonance \ion{Fe}{2}~$^*$ transitions and spatially
extended \ion{Mg}{2} emission which they interpret as originating from
photons scattered off a galactic-scale wind.  They estimate that the
wind extends to at least 5\,kpc from the galaxy.  Line emission is
also observed for $z \sim 3$ LBGs via the resonant \lya\ transition
and fine-strcuture \ion{Si}{2}$^*$ transitions \citep{cb58,shapley}.
\cite{rubin+10a} emphasize that a comprehensive analysis of these data
(e.g.\ via integral-field-unit observations) could yield unique
diagnostics on the morphology and density of the outflow, eventually
yielding tigheter constrains on the energetics of the flow.  
The crucial piece missing from this path is a radiative transfer
treatment that properly ...

In this paper, we take the first steps toward modeling the absorption
and emission properties of cool gas outflows.  Using Monte Carlo
radiative transfer technqiues, we study the nature of \ion{Mg}{2} and
\ion{Fe}{2} absorption and emission for winds with a range of
properties.  Although the winds are idealized, the results frequently
contradict our initial intuition and caution one [word] on the
challenges of converting the observables to (even crude) physical
constraints.  [Add another line or two]

[Outline of paper]

[Point out IR emission from winds locally implies dust]
